Synthesizing Symmetric Lenses

### 2018-07-12 papers/SymmetricLenses/main.tex
merge kathleens changes in

% begin evaluation
\section{Evaluation}
\label{sec:evaluation}
We have implemented our symmetric DNF lenses and the synthesis of them as an
extension to Boomerang. We added symmetric lens combinators to Boomerang, and
implemented all of Boomerang's asymmetric lens combinators 
using a combination of the symmetric lens combinators presented in this
paper and symmetric versions of the asymmetric theory extensions
already present in Boomerang, like matching lenses and quotient lenses.
We also integrated our synthesis engine into Boomerang. It allows users to write synthesis tasks
alongside lens combinators, use those synthesis results in the manual writing of
new lenses, and use previously defined lenses in the synthesis of new lenses.

In this evaluation, we aim to answer three primary questions:
\begin{enumerate}
\item Is our synthesis procedure efficient enough to be used in everyday
  development?

\item How much slower is our tool on bijective lens synthesis benchmarks than
  the existing Optician system customized for bijective lenses?

\item Which heuristics are useful for reducing the ``specification
  overhead'' for users? In particular, which heuristics are useful in
  synthesizing the desired lens, instead of merely a lens that satisfies the
  specification?
\end{enumerate}

\subsection{Benchmark Suite}
Our benchmarks are drawn from three different sources.
\begin{enumerate}
\item We adapted 8 data cleaning benchmarks from Flash Fill, though, of course, our tool produces bidirectional transformations rather than one-way transformers like Flash Fill.
\item We adapted 29 benchmarks from Augeas. These benchmarks turn the
  ad hoc data present in Linux configuration files into a structured, dictionary
  representation.
\item We used 11 benchmarks from synchronization tasks between a variety of
  formats derived from real-world examples and the bidirectional programming
  literature. These tasks range from synchronizing REST and JSON web resource
  descriptions to synchronizing BibTeX and EndNote citation descriptions.
\end{enumerate}

A realistic benchmarks suite is critical in answering the first research
question. To determine whether our algorithm is fast enough to be used in
everyday development, we need to evaluate whether we can solve real-world
problems efficiently. All three of our benchmark sources test real life
problems: the data cleaning tasks presented in Flash Fill are either derived
from online help forums or taken from the Excel product team; Augeas is a
utility that transforms Linux configuration files into a manageable dictionary
form; the many of the remaining benchmarks are from synchronizing real-life data
formats, like BibTeX and EndNote.

@@

To begin, we measure the run time of our algorithm in 2 modes:

@@

\begin{figure}
  \includegraphics{generated-graphs/times}
  \caption{Number of benchmarks that can be solved by a given algorithm in a
    given amount of time. \SSOpt{} is the full symmetric synthesis algorithm.
    \SSNCOpt{} is the symmetric synthesis algorithm without using a library of
    existing lenses. Our symmetric synthesis algorithm is able to complete all
    benchmarks in our benchmark suite in under 30 seconds, and is able to
    complete a majority of the benchmarks without using compositional
    synthesis.}
  \label{fig:times}
\end{figure}

For each benchmark in our benchmark suite, we ran every mode on it with a
timeout of 60 seconds, and averaged the result over 5 runs.
Figure~\ref{fig:times} summarizes the results of these tests. We find that our
algorithm is able to synthesize all of the benchmarks in our benchmark suite in
under 30 seconds. Without compositional synthesis, the synthesis algorithm is
able to solve 26 problem instances. Recall from
Section~\ref{subsec:optimizations} that users may tune the termination metric.
Success on this set of benchmarks requires we tune that termination codition
(e.g. for some benchmarks, allowing the system flexibility to spend more time
considering more of the search space). In
Section~\ref{subsec:effect-of-heuristics}, we will consider the effect of fixing
the termination metric. In practice, we found customizing the termination metric
to be quite easy: if manual inspection of the lens showed there were too few
bijections, we would increase the termination parameter. If synthesis took too
long, we would decrease it.

@@

To compare to the existing bijective synthesis algorithm, we run our symmetric
synthesis algorithm on the original Optician benchmark suite. 4 of these
benchmarks had to be slightly altered to work with our new system. We altered
these benchmarks by either providing additional examples or tuning termination parameters.
Without these alterations, the symmetric synthesis yielded a lens that fit
the specification but that was not the desired lens.

To perform this comparison, we synthesized lenses in 2 modes:

@@

For each benchmark in our benchmark suite, we ran every mode on it with a
timeout of 60 seconds, and averaged the result over 5 runs.
Figure~\ref{fig:times_bijective} summarizes the results of these tests. On
average, \SSOpt{} took 1.3 times (0.5 seconds) longer to complete than
\BSOpt{}. The slowest completed benchmark for both synthesis algorithms is
\texttt{xml\_to\_augeas.boom}, a benchmark that converts arbitrary XML up to
depth 3 into a serialized version of the structured dictionary representation
used in Augeas. This benchmark takes 18.9 seconds for the symmetric synthesis
algorithm to complete, and 9.3 seconds the bijective synthesis algorithm to
complete.

@@

\subsection{The Effects of Our Heuristics}
\label{subsec:effect-of-heuristics}
We evaluate the usefulness of (1) our information-theoretic metric, and (2) our
termination heuristic.  To this end, we run our program in several different
modes:\\
\begin{tabulary}{\linewidth}{rL}
  \AnyOpt{}: & We synthesize lenses without use of the information-theoretic
  preference metric.  Once \GreedySynth finds a lens, the 
  synthesis procedure returns that lens.\\
  \FLOpt{}: &  The first time \GreedySynth finds a lens, the synthesis procedure
              returns that lens. However, \GreedySynth uses the
              information-theoretic preference metric to choose a lens.\\ 
  \CCOpt{}: & We replace our information-theoretic cost metric with one where
              the cost of the lens is the number of disonnects plus the number
              of merges.\\
  \ZTOpt{}: & \ZTOpt{} runs the algorithm with the termination parameter
              universally set to 0 (default). \\
  \TFTOpt{}: & \TFTOpt{} runs the algorithm with the
               termination parameter universally set to 25. \\
  \NTFTOpt{}: & \NTFTOpt{} runs the algorithm with the
                termination parameter universally set to -25. \\
\end{tabulary}

@@

We experimented with the \CCOpt{} mode to determine whether the complexity of the information-theoretic measure was really needed. Related work on string transformations has often used simpler measures such as ``avoid constants'' which align with, but are simpler than our measures. The \CCOpt{} mode is an example of such a simple measure---it operates by counting disconnects, which put a complete stop to information transfer, like constants, and merges, which eliminate the information in a union. Figure~\ref{fig:metric} summarizes the result of this experiment. We find that
our metrics are very important for finding the correct lens. Our
information-theoretic metric is critical for finding the correct lens. Indeed,
only 10 of our benchmarks succeeded when running in \CCOpt{} mode.

Our termination condition is also quite important. When running in \FLOpt{},
only 5 lenses are found. The first cluster in which a lens is found is
rarely the correct cluster. However, our termination condition is not perfect.
Sometimes users must tell the algorithm to ``keep going'' or ``stop early.''
Without termination parameters, our algorithm finds the correct lens for 36 of
our 48 benchmarks, 9 of them need help in the form of a termination parameter
for finding the correct lens. That being said, our default parameter finds a
good middle-ground. Both \TFTOpt{} and \NTFTOpt{} synthesized the correct lens
less often than \ZTOpt.  \TFTOpt{} necessitates too good a lens, so the search
procedure keeps spinning, and \NTFTOpt{} terminates early, not finding the
correct lens.
% end evaluation

>>>>

% begin evaluation
\section{Evaluation}
\label{sec:evaluation}
We have implemented symmetric DNF lenses as an extension to
Boomerang. Specifically, we added symmetric lens combinators to
Boomerang, and implemented all of Boomerang's asymmetric lens
combinators  using a combination of the symmetric lens combinators presented in this
paper and symmetric versions of the asymmetric extensions
already present in Boomerang, like matching lenses and quotient lenses.
We also integrated our synthesis engine into Boomerang. This
integration allows users to write synthesis tasks 
alongside lens combinators, incorporate synthesis results into 
manually-written lenses, and reference previously defined lenses in
the synthesis of new lenses.

In this evaluation, we aim to answer three primary questions:
\begin{enumerate}
\item Does the synthesis algorithm find the correct lens?
\item Is the synthesis procedure efficient enough to be used in everyday
  development?
\item How much slower is our tool on bijective lens synthesis benchmarks than
  the existing Optician system customized for bijective lenses?

\item Which heuristics are useful
for reducing the ``specification 
  overhead'' for users? In particular, which heuristics are useful in
  synthesizing the desired lens instead of merely a lens that satisfies the
  specification?
\end{enumerate}

\subsection{Benchmark Suite}
Our benchmarks are drawn from three different sources.
\begin{enumerate}
\item We adapted 8 data cleaning benchmarks from Flash Fill. Note
that our tool produces bidirectional transformations rather
than one-way transformers like Flash Fill. 
\item We adapted 29 benchmarks from Augeas. These benchmarks turn the
  ad hoc data present in Linux configuration files into a structured dictionary
  representation.
\item We created 11 additional benchmarks 
  derived from real-world examples and/or the bidirectional programming
  literature. These tasks range from synchronizing REST and JSON web resource
  descriptions to synchronizing BibTeX and EndNote citation descriptions.
\end{enumerate}
Each benchmark comprises a source and target regular expression, a
collection of examples, and an optional termination condition
parameter. 
%%A realistic benchmarks suite is critical in answering the first research
%%question. To determine whether our algorithm is fast enough to be used in
%%everyday development, we need to evaluate whether we can solve real-world
%%problems efficiently.

All three sources for our benchmarks provide realistic problems
suitable for answering the research questions:
Flash Fill  data cleaning tasks are either derived from online help
forums or taken from the Excel product team;  
Augeas is a utility that transforms Linux configuration files into a manageable dictionary
form; and
many of the remaining benchmarks arise from synchronizing real-world
data formats such as BibTeX and EndNote.

\subsection{Lens correctness}
To determine whether the synthesized lenses were the desired ones,
we ran the symmetric synthesis algorithm on the 48 benchmarks and then
manually inspected the generated lenses.  In all 48 cases, the
synthesized benchmark was correct.

@@

We measured the run time of our algorithm in two modes:

@@

\begin{figure}
  \includegraphics{generated-graphs/times}
  \caption{Number of benchmarks that can be solved by a given algorithm in a
    given amount of time. \SSOpt{} is the full symmetric synthesis algorithm.
    \SSNCOpt{} is the symmetric synthesis algorithm without using a library of
    existing lenses. The symmetric synthesis algorithm is able to complete all
    benchmarks in under 30 seconds elapsed total time. In addition, it is able to
    complete a majority of the benchmarks in that timeframe without
  using compositional 
    synthesis.}
  \label{fig:times}
\end{figure}

For each benchmark in our benchmark suite, we ran it in both modes with a
timeout of 60 seconds, and averaged the result over 5 runs.
Figure~\ref{fig:times} summarizes the results of these tests. We find that our
algorithm is able to synthesize all of the benchmarks in our benchmark suite in
under 30 seconds. Without compositional synthesis, the synthesis algorithm is
able to solve 26 problem instances. Recall from
Section~\ref{subsec:optimizations} that users may tune the termination metric.
Success on this set of benchmarks requires we tune that termination condition
(e.g. for some benchmarks, allowing the system flexibility to spend more time
considering more of the search space). In
Section~\ref{subsec:effect-of-heuristics}, we will consider the effect of fixing
the termination metric. In practice, we found customizing the termination metric
to be quite easy: if manual inspection of the lens showed there were too few
bijections, we would increase the termination parameter. If synthesis took too
long, we would decrease it.

@@

To compare to the existing bijective synthesis algorithm, we run our symmetric
synthesis algorithm on the original Optician benchmarks, comprised of
39 bijective synthesis tasks. We had to slightly alter four of these
benchmarks, either by providing additional examples or by tuning termination parameters.
Without these alterations, symmetric synthesis yielded a lens that fit
the specification but that was not the desired lens.

To perform this comparison, we synthesized lenses in two modes:

@@

For each benchmark, we ran it in both modes with a
timeout of 60 seconds and averaged the result over 5 runs.
Figure~\ref{fig:times_bijective} summarizes the results of these tests. On
average, \SSOpt{} took 1.3 times (0.5 seconds) longer to complete than
\BSOpt{}. The slowest completed benchmark for both synthesis algorithms is
\texttt{xml\_to\_augeas.boom}, a benchmark that converts arbitrary XML up to
depth 3 into a serialized version of the structured dictionary representation
used in Augeas. This benchmark takes 18.9 seconds for the symmetric synthesis
algorithm to complete, and 9.3 seconds the bijective synthesis algorithm to
complete.

@@

\subsection{The Effects of Our Heuristics}
\label{subsec:effect-of-heuristics}
We evaluate the usefulness of (1) our information-theoretic metric, and (2) our
termination heuristic.  To this end, we run our program in several different modes:\\[1ex]
\begin{tabulary}{\linewidth}{rL}
  \AnyOpt{}: & We ignore the information-theoretic preference metric entirely,
  returning the first lens \GreedySynth finds without using this information. \\ 
  \FLOpt{}: &  \GreedySynth uses the information-theoretic preference
  metric to find a lens, but only the original pair of stochastic
  regular expressions are considered as end-points.  \\
  \CCOpt{}: & We replace our information-theoretic cost metric with one where
              the cost of the lens is the number of disonnects plus the number
              of merges.\\
  \ZTOpt{}: & \ZTOpt{} runs the algorithm with the termination parameter
              universally set to 0 (default). \\
  \TFTOpt{}: & \TFTOpt{} runs the algorithm with the
               termination parameter universally set to 25. \\
  \NTFTOpt{}: & \NTFTOpt{} runs the algorithm with the
                termination parameter universally set to -25. \\
\end{tabulary}

@@

We experimented with the \CCOpt{} mode to determine whether the
complexity of the information-theoretic measure was really
needed. Related work on string transformations has often used simpler
measures such as ``avoid constants'' which align with, but are simpler
than our measures. The \CCOpt{} mode is an example of such a simple
measure---it operates by counting disconnects, which put a complete
stop to information transfer, and merges, which
eliminate the information in a union.

Figure~\ref{fig:metric}
summarizes the result of these experiments. The data reveal that 
the information-theoretic metric is critical for finding the correct
lens: Only 10 of the benchmarks succeeded when running in \CCOpt{} mode.
The termination condition is also quite important. When running
in \FLOpt{} mode, the algorithm only discovers 5 lenses, which shows
that the first cluster that contains a satisfying lens is rarely the
correct cluster.  However, our termination condition is not perfect.
Sometimes users must tell the algorithm to ``keep going'' or ``stop early.''
Without termination parameters, our algorithm finds the correct lens for 36 of
our 48 benchmarks.  Nine of them need help in the form of a termination parameter
for finding the correct lens. That being said, our default parameter finds a
good middle-ground. Both \TFTOpt{} and \NTFTOpt{} synthesized the correct lens
less often than \ZTOpt.  \TFTOpt{} demands too good a lens, so the search
procedure keeps spinning, while \NTFTOpt{} terminates early, not
always trying hard enough to find the correct lens.
% end evaluation
